Open source landscape ecology tools

### 2020-07-26 paper/paper.Rmd
adds references to data representation section

The most prominent package allowing for raster data representation is *raster*.
A *raster* successor aimed at the simpler interface and improved performance, *terra*, is now being developed, however, it could take several years for this package to be adopted by users.
Alternatively, *stars* package can be used to read and process raster data.
It is interconnected with the *sf* package, and beside raster support, it provides tools for spatial data cubes.

The *sp* package was a standard for vector data representation for more than 10 years.
As of 2020, more than 500 R packages directly depends or imports *sp*.
However, it is now in a maintenance mode only with a recommended successor - the *sf* package.
<!-- terra new vector representations -->

>>>>

The most prominent package allowing for raster data representation is *raster* \cite{Hijmans2019}.
A *raster* successor aimed at the simpler interface and improved performance, *terra*, is now being developed \cite{R-terra}, however, it could take several years for this package to be adopted by users.
Alternatively, *stars* package can be used to read and process raster data \cite{Pebesma2019}.
It is interconnected with the *sf* package, and beside raster support, it provides tools for spatial data cubes.

The *sp* package was a standard for vector data representation for more than 10 years \cite{Pebesma2005,Bivand2013}.
As of 2020, more than 500 R packages directly depends or imports *sp*.
However, it is now in a maintenance mode only with a recommended successor - the *sf* package.
<!-- terra new vector representations -->
